---
title: "Gold Forecast Prophet"
author: "Amin"
date: "11/24/2022"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
library(fpp2) 
library(zoo)
library(tidyr)
library(ggplot2)
library(readr)
library(dplyr)
library(rvest)
library(Quandl)
library(plotly)
library(ggplot2)
library(DT)
library(gtrendsR)
library(tidytext)
library(stringr)
library(lubridate)
library(prophet)
library(imputeTS)
library(corrplot)
#library(tidyverse)
library(tidymodels)
#install.packages("tidytext")
#install.packages("imputeTS")
```


## Load Data
```{r}
# Load Data

gold <- read_csv("/Users/datascience/Desktop/Time Series Data Science/Time Series Project/Update/Gold.csv", 
    col_types = cols(Date = col_date(format = "%m/%d/%Y")))

unemp <- read_csv("/Users/datascience/Desktop/Time Series Data Science/Time Series Project/Update/UNRATE.csv", show_col_types = FALSE)

M2 <- read_csv("/Users/datascience/Desktop/Time Series Data Science/Time Series Project/Update/WM2NS.csv", show_col_types = FALSE)

DJ <- read_csv("/Users/datascience/Desktop/Time Series Data Science/Time Series Project/Update/DJIA.csv",  show_col_types = FALSE)                

fed_funds <- read_csv("/Users/datascience/Desktop/Time Series Data Science/Time Series Project/Update/DFF.csv", show_col_types = FALSE)

silver <- Quandl('LBMA/SILVER')

dollar_index <- read_csv("/Users/datascience/Desktop/Time Series Data Science/Time Series Project/Update/US Dollar Index Historical Data.csv", col_types = cols(Date = col_date(format = "%m/%d/%Y")))

Google <- read_csv("/Users/datascience/Desktop/Time Series Data Science/Time Series Project/Update/Gold_Trend_hits.csv", show_col_types = FALSE)
```

## Create Candlestick/volume graph of gold
```{r, warning=FALSE}
#edit the column names in data frames 
Price_plot <- gold %>%
  plot_ly(x = ~Date,
          type = "candlestick", 
          open = ~Open, 
          close = ~`Close/Last`, 
          high = ~High,
          low = ~Low,
          name = "price") %>%
  layout(
    xaxis = list(
      rangeselector = list(
        buttons = list(
          list(
            count = 3,
            label = "3 mo",
            step = "month",
            stepmode = "backward"),
          list(
            count = 6,
            label = "6 mo",
            step = "month",
            stepmode = "backward"),
          list(
            count = 1,
            label = "1 yr",
            step = "year",
            stepmode = "backward"),
          list(
            count = 2,
            label = "2 yr",
            step = "year",
            stepmode = "backward"),
          list(
            count = 3,
            label = "3 yr",
            step = "year",
            stepmode = "backward"),
          list(
            count = 5,
            label = "5 yr",
            step = "year",
            stepmode = "backward"),
          list(step = "all"))),
      
      rangeslider = list(visible = FALSE)),
         yaxis = list(title = "Price ($)",
                      showgrid = TRUE,
                      showticklabels = TRUE))
Volume <- select(gold, Date, Volume)
Volume$Date <- as.Date(Volume$Date , format = "%m/%d/%y")
Volume$Vol <- as.numeric(as.character(Volume$Volume)) / 1000
Volume_plot <- Volume %>%
  plot_ly(x=~Date, y=~Vol, type='bar', name = "Volume") %>%
  layout(yaxis = list(title = "Volume (Units of Thousand)"))
plot <- subplot(Price_plot, Volume_plot, heights = c(0.7,0.3), nrows=2,
             shareX = TRUE, titleY = TRUE) %>%
  layout(title = 'GC:CMX')
plot
```

## Adjust data types where needed 
```{r}
#change data type from chr to date in the gold data set
gold$Date <- as.Date(gold$Date, format = "%m/%d/%y")
gold$DATE <- gold$Date
#change the datatype to date in the DJIA data set
DJ$DATE <- as.Date(DJ$DATE, format = "%m/%d/%y")
#indicate columns to keep in gold data set
keep_cols <-  c("Date","Close/Last", 'Volume')
gold <- gold[keep_cols] 
gold <- gold %>%
  rename(Gold_Close = 'Close/Last', Gold_Volume ='Volume')
```
## Create the daily M2 rate data frame (in billions)
```{r}
#create all dates from the date column in the data frame
all_dates <- M2 %>%select(DATE) %>% 
  complete(DATE=  seq.Date(min(DATE), max(DATE), by="day"),
           ) %>% mutate(TIME=paste(year(DATE),str_pad(month(DATE), 2, pad = "0"),sep = "-"))
#join the all_dates object back onto original datafarme
M2_daily <- left_join(all_dates,M2,by="DATE")
#fill the NA values from the join with the first value from the dataset
M2_daily <- na_locf(M2_daily, option = "locf")
keep_cols <- c("DATE", "WM2NS")
M2_daily <- M2_daily[keep_cols]
M2_daily <- M2_daily %>%
  select(c("DATE","WM2NS")) %>%
  rename(Date = DATE)

```

## Create the daily unemployment rate data frame
```{r}
#create all dates from the date column in the dataframe
all_dates <- unemp %>%select(DATE) %>% 
  complete(DATE=  seq.Date(min(DATE), max(DATE), by="day"),
           ) %>% mutate(TIME=paste(year(DATE),str_pad(month(DATE), 2, pad = "0"),sep = "-"))
#join the all_dates object back onto original datafarme
unemp_daily <- left_join(all_dates,unemp,by="DATE")
#fill the NA values from the join with the first value from the dataset
unemp_daily <- na_locf(unemp_daily,  option = "locf")
keep_cols <- c("DATE", "UNRATE")
unemp_daily <- unemp_daily[keep_cols]
unemp_daily <- unemp_daily %>%
  select(c("DATE","UNRATE")) %>%
  rename(Date = DATE)
```

## Create the daily federal funds data frame
```{r}
#create all dates from the date column in the data frame
all_dates <- fed_funds %>%select(observation_date) %>% 
  complete(observation_date =  seq.Date(min(observation_date), max(observation_date), by="day"),
           ) %>% mutate(TIME=paste(year(observation_date),str_pad(month(observation_date), 2, pad = "0"),sep = "-"))
#join the all_dates object back onto original dataframe
ff_daily <- left_join(all_dates,fed_funds,by="observation_date")
#fill the NA values from the join with the first value from the dataset
ff_daily <- na_locf(ff_daily, option = "locf")
keep_cols <- c("observation_date", "DFF")
ff_daily <- ff_daily[keep_cols]
ff_daily <- ff_daily %>%
  select(c("observation_date","DFF")) %>%
  rename(Date = observation_date)
```
## Create the daily Dow Jones IA data frame
```{r}

#create all dates from the date column in the data frame
all_dates <- DJ %>%select(DATE) %>% 
  complete(DATE =  seq.Date(min(DATE), max(DATE), by="day"),
           ) %>% mutate(TIME=paste(year(DATE),str_pad(month(DATE), 2, pad = "0"),sep = "-"))
#join the all_dates object back onto original datafarme
dj_daily <- left_join(all_dates, DJ ,by="DATE")

#fill the NA values from the join with the first value from the dataset
dj_daily <- na_locf(dj_daily,  option = "locf")
keep_cols <- c("DATE", "DJIA")
dj_daily <- dj_daily[keep_cols]
dj_daily <- dj_daily %>%
  select(c("DATE", "DJIA")) %>%
  rename(Date = DATE)

```

## Create the daily Silver Price data frame
```{r}
# add in Silver
silver <- silver[,c('Date', 'USD')]
silver <- silver %>%
  rename(Silver_Close = 'USD')
# Join Silver and Gold Data
Metals <- inner_join(silver, gold, by="Date")
```

## Create the Dollar Index Price data frame
```{r}
# Add in Dollar Index
dollar_index <- dollar_index %>%
  select(c("Date","Price")) %>%
  rename(DXY = Price)
```

## Google trends to find hits on Gold worlwide
```{r}
#trends <- gtrends(keyword = "Gold", onlyInterest = TRUE, time = "today+5-y")

#create all dates from the date column in the data frame
all_dates <- Google %>%select(Date) %>% 
  complete(Date =  seq.Date(min(Date), max(Date), by="day"),
           ) %>% mutate(TIME=paste(year(Date),str_pad(month(Date), 2, pad = "0"),sep = "-"))
#join the all_dates object back onto original datafarme
Google_daily <- left_join(all_dates, Google ,by="Date")
#fill the NA values from the join with the first value from the dataset
Google_daily <- na_locf(Google_daily,  option = "locf")

# Create Daily Data Fame
keep_cols <- c("Date", "Gold: (Worldwide)")
Google_daily <- Google_daily[keep_cols]
Google_daily <- Google_daily %>%
  select(c("Date", "Gold: (Worldwide)")) %>%
  rename(hits = 'Gold: (Worldwide)')
Google_daily %>%  
  plot_ly(type='scatter',x=~Date, y=~hits, mode = 'lines', name = "Google Search Trends") %>%
  layout(title = paste0("Interest over Time: ", "Gold"), yaxis = list(title = "hits"))
trends_daily <- Google_daily
# Hits verse Gold 
trends_daily %>%
  left_join(gold, by = "Date") %>%
  select(one_of(c("Date", "hits", "Gold_Close"))) %>%
  drop_na() %>%
  ggplot(aes(hits, Gold_Close)) + geom_point(color="red") + geom_smooth(method = 'loess') +
  labs(title =paste0("Gold",": Relationship between World Interest (Hits) and Close Price (Gold)"))

```
## Merge data frames and create the final dataframe for EDA
```{r}
#merge the dataframes together on dates
full_df <- left_join(Metals, dollar_index, by="Date")
full_df <- left_join(full_df, unemp_daily, by="Date")
full_df <- left_join(full_df, ff_daily, by="Date")
full_df <- left_join(full_df, M2_daily, by="Date")
full_df <- left_join(full_df, Google_daily, by="Date")
full_df <- left_join(full_df, dj_daily, by="Date")
full_df <- left_join(full_df, trends_daily, by="Date")
full_df$Gold_Volume <- na_ma(full_df$Gold_Volume, k=1, weighting = "simple")
keep_cols <- c("Date", "Silver_Close", "Gold_Close","Gold_Volume", "DXY", "UNRATE", "DFF", "WM2NS" ,"hits.x",     "DJIA")     
full_df <- full_df[keep_cols]
full_df <- full_df %>%
  rename(hits = hits.x)
```

```{r}
# Check for any NA Values
cbind(
   lapply(
     lapply(full_df, is.na)
     , sum)
   )
```
## Subset the data frame with values from 2018 forward
```{r}
full_df <- full_df %>% arrange(ymd(full_df$Date))
full_df <- full_df[full_df$Date >= "2018-01-01",]
#Replace NA with the most recent value 
full_df$WM2NS <- zoo::na.fill(full_df$WM2NS, "extend")
full_df$hits <- zoo::na.fill(full_df$hits, "extend")
full_df$UNRATE <- zoo::na.fill(full_df$UNRATE, "extend")
tail(full_df)
```
```{r}
# Check for any NA Values
cbind(
   lapply(
     lapply(full_df, is.na)
     , sum)
   )
```

## Create the correlation matrix for EDA 
```{r}
corr_fields <- c("Gold_Close", "Silver_Close","Gold_Volume", "DXY", "UNRATE", "DFF", "WM2NS" ,"hits", "DJIA")     
full_df$Gold_Volume <- as.numeric(full_df$Gold_Volume)
full_df$WM2NS <- as.numeric(full_df$WM2NS)
full_df$hits <- as.numeric(full_df$hits)

full_df$Gold_Volume <- na_ma(full_df$Gold_Volume, k=1, weighting = "simple")
full_df$WM2NS <- na_ma(full_df$WM2NS, k=1, weighting = "simple")
full_df$hits <- na_ma(full_df$hits, k=1, weighting = "simple")
```

```{r}
corr_df <- full_df[corr_fields]
corr_matrix = cor(corr_df)
#display first row of correlation matrix
corr_matrix[,1]
```
```{r}
#develop the corrplot correlation matrix
corrplot(corr_matrix)
```
Silver Close, DFF, WM2NS, hits, and DIJA are strongly correlated with gold and therefore will be used as external regressors.
## Normalize variables for EDA
```{r}
#normalize variables to view scaled relationships and add into full_df
full_df_Norm <- full_df
full_df_Norm[-1] <- lapply(full_df_Norm[-1], scale)
```

## EDA with non-normalized varialbes
````{r}
#scatterplot of gold & each external variable

ggplot(data = full_df) +
  geom_point(mapping = aes(x = Gold_Close,
                            y = UNRATE, color = Date)) + 
  ggtitle("Scatterplot of Gold Close Price and Unemployment Rate (Date Color)")

ggplot(data = full_df) +
  geom_point(mapping = aes(x = Gold_Close,
                            y = DFF, color = Date)) + 
  ggtitle("Scatterplot of Gold Close Price and Federal Funds Rate (Date Color)")

ggplot(data = full_df) +
  geom_point(mapping = aes(x = Gold_Close,
                            y = WM2NS, color = Date)) + 
  ggtitle("Scatterplot of Gold Close Price and M2 Money Supply (Date Color)")

ggplot(data = full_df) +
  geom_point(mapping = aes(x = Gold_Close,
                            y = DJIA, color = Date)) + 
  ggtitle("Scatterplot of Gold Close Price and DJIA (Date Color)")

ggplot(data = full_df) +
  geom_point(mapping = aes(x = Gold_Close,
                            y = Gold_Volume, color = Date)) + 
  ggtitle("Scatterplot of Gold Close Price and Gold Volume (Date Color)")

ggplot(data = full_df) +
  geom_point(mapping = aes(x = Gold_Close,
                            y = Silver_Close, color = Date)) + 
  ggtitle("Scatterplot of Gold_Close and Silver_Close (Date Color)")

ggplot(data = full_df) +
  geom_point(mapping = aes(x = Gold_Close,
                            y = hits, color = Date)) + 
  ggtitle("Scatterplot of Gold_Close and Hits (Date Color)")
```
## Create dataframe for Normalize EDA (on same axis)
```{r}
#create time series objects for each of the normalized data elements 
gold_ts_norm <- ts(full_df_Norm$Gold_Close, start = c(2018, 01), frequency = 252)
unemp_ts_norm <- ts(full_df_Norm$UNRATE, start = c(2018, 01), frequency = 252)
ff_ts_norm <- ts(full_df_Norm$DFF, start = c(2018, 01), frequency = 252)
M2_ts_norm <- ts(full_df_Norm$WM2NS, start = c(2018, 01), frequency = 252)
dj_ts_norm <- ts(full_df_Norm$DJIA, start = c(2018, 01), frequency = 252)
silver_ts_norm <- ts(full_df_Norm$Silver_Close, start = c(2018, 01), frequency = 252)
hits_ts_norm <- ts(full_df_Norm$hits, start = c(2018, 01), frequency = 252)
gold_vol_ts_norm <- ts(full_df_Norm$Gold_Volume, start = c(2018, 01), frequency = 252)
```

## Create time series plots of all factors
```{r}
#line chart of all indicators
autoplot(gold_ts_norm, main = "5-Year Trend of Gold and Unemployment Rate", xlab = "Date") + 
  autolayer(unemp_ts_norm) + theme_classic()

autoplot(gold_ts_norm, main = "5-Year Trend of Gold and Federal Funds Rate", xlab = "Date") + 
  autolayer(ff_ts_norm) + theme_classic()

autoplot(gold_ts_norm, main = "5-Year Trend of Gold Close and M2", xlab = "Date") + 
  autolayer(M2_ts_norm) + theme_classic()

autoplot(gold_ts_norm, main = "5-Year Trend of Gold and DJIA", xlab = "Date") + 
  autolayer(dj_ts_norm) + theme_classic()

autoplot(gold_ts_norm, main = "5-Year Trend of Gold and Silver Close", xlab = "Date") + 
  autolayer(silver_ts_norm) + theme_classic()

autoplot(gold_ts_norm, main = "5-Year Trend of Gold and Hits", xlab = "Date") + 
  autolayer(hits_ts_norm) + theme_classic()

autoplot(gold_ts_norm, main = "5-Year Trend of Gold and Gold Volume", xlab = "Date") + 
  autolayer(gold_vol_ts_norm) + theme_classic()



```

#### Boxplot to identify outlier variables 
```{r}
#boxplot of all varialbes  
ggplot(corr_df, aes(x=Gold_Close)) +
  geom_boxplot() + ggtitle("Gold Boxplot") + xlab("Gold Closing Price")

ggplot(corr_df, aes(x=UNRATE)) +
  geom_boxplot() + ggtitle("Unemployement Rate Boxplot") + xlab("Unemployement Rate")

ggplot(corr_df, aes(x=DFF)) +
  geom_boxplot() + ggtitle("Federal Funds Rate Boxplot") + xlab("Federal Funds Rate")

ggplot(corr_df, aes(x=WM2NS)) +
  geom_boxplot() + ggtitle("M2 Money Supply Boxplot") + xlab("M2 Total Money Supply")

ggplot(corr_df, aes(x=DJIA)) +
  geom_boxplot() + ggtitle("DJIA Boxplot") + xlab("Dow Jones Industrial Average")

ggplot(corr_df, aes(x=Gold_Volume)) +
  geom_boxplot() + ggtitle("Gold Volume Boxplot") + xlab("Gold Volume")

ggplot(corr_df, aes(x=Silver_Close)) +
  geom_boxplot() + ggtitle("Silver Close Boxplot") + xlab("Silver Close")

ggplot(corr_df, aes(x=hits)) +
  geom_boxplot() + ggtitle("Hits Boxplot") + xlab("Hits")
```

## Full entire Dataset

```{r}
FULL_df  <- left_join(full_df, full_df_Norm, by="Date")
```

## Partition the data into training and testing
```{r}
#create the time series object for the FULL_df
ts_FULL_df <- ts(data=FULL_df$Gold_Close.x, start = c(2018,01), frequency = 252)
train_df <- window(ts_FULL_df, end = c(2022,150)) #161 is end of Aug
test_df <-  window(ts_FULL_df, start = c(2022,151))
#validate the train and test layers connect
autoplot(train_df) + autolayer(test_df) +
  #coord_cartesian(xlim = c(2022,2023)) +
  labs(title = "2022 - 2023 Gold Close Price (Train & Test)",
       x = "Time", y = "Gold Close Price")
test_df
```



## Function to find local min and max for buy/sell signals 
```{r}
 # Locate Local Min and Max for buy/sell signals
locate_xtrem <- function (x, last = FALSE)
{
  # use rle to deal with duplicates
  x_rle <- rle(x)

  # force the first value to be identified as an extrema
  first_value <- x_rle$values[1] - x_rle$values[2]

  #
  # ! NOTE: with this method, last value will be considered as an extrema
  diff_sign_rle <- c(first_value, diff(x_rle$values)) %>% sign() %>% rle()

  # this vector will be used to get the initial positions
  diff_idx <- cumsum(diff_sign_rle$lengths)

  # find min and max
  diff_min <- diff_idx[diff_sign_rle$values < 0]
  diff_max <- diff_idx[diff_sign_rle$values > 0]

  # get the min and max indexes in the original series
  x_idx <- cumsum(x_rle$lengths)
  if (last) {
    min <- x_idx[diff_min]
    max <- x_idx[diff_max]
  } else {
    min <- x_idx[diff_min] - x_rle$lengths[diff_min] + 1
    max <- x_idx[diff_max] - x_rle$lengths[diff_max] + 1
  }
  # just get number of occurences
  min_nb <- x_rle$lengths[diff_min]
  max_nb <- x_rle$lengths[diff_max]

  # format the result as a tibble
  bind_rows(
    tibble(Idx = min, Values = x[min], NB = min_nb, Status = "min"),
    tibble(Idx = max, Values = x[max], NB = max_nb, Status = "max")) %>%
    arrange(.data$Idx) %>%
    mutate(Last = last) %>%
    mutate_at(vars(.data$Idx, .data$NB), as.integer)
}

```

## GOLD Prophet FORECAST BASELINE (No External Regressors)

```{r}
# Partition Data frame
prophet.Train <- subset(FULL_df, Date < as.Date("2022-11-08"))
Prophet.Test <- subset(FULL_df, Date >= as.Date("2022-11-17"))

# Set training col for prophet forecast
prophet.Train <- prophet.Train %>%
  select(c("Date","Gold_Close.x")) %>%
  rename(ds = Date, y = Gold_Close.x)

# Prophet Predictions
Prophet <- prophet(prophet.Train, interval.width = 0.95)
future <- make_future_dataframe(Prophet, periods = 7) %>% filter(!wday(ds) %in% c(1,7)) #account for regular gaps on weekends
Prophet_Forecast_base <- predict(Prophet, future) 

# Grab necessary variables
Forecast_subset <- Prophet_Forecast_base %>%
  select(c('ds','yhat','yhat_lower','yhat_upper')) %>%
  rename(Date = ds, ClosePrice = yhat , ClosePrice_lower = yhat_lower, ClosePrice_upper = yhat_upper)

# Put into Dataframe
datatable(Forecast_subset[c('Date','ClosePrice','ClosePrice_lower','ClosePrice_upper')])

# Return prediction results from prophet forecast 
Prophet_Results_base <- Prophet_Forecast_base %>%
  select(c("ds","yhat")) %>%
  rename(Date = ds, Close_Prediction = yhat)
Prophet_Results_base$Date <- as.Date(Prophet_Results_base$Date , format = "%m/%d/%y")
Prophet_Results_base <- subset(Prophet_Results_base, Date >= as.Date("2022-11-08"))

# Grab actual gold data
Gold_Results <- gold %>%
  select(c("Date","Gold_Close"))  %>%
  rename(Close_Actual = Gold_Close)
 
# Join prediction and actual results for comparison
results <- left_join(Prophet_Results_base, Gold_Results,  by="Date")
results

# Calculate RMSE 
RMSE <- sqrt(mean((results$Close_Actual - results$Close_Prediction)^2))
RMSE 

```

## Plot Baseline Prophet Forecast
```{r}
# Plot prophet forecast
plot(Prophet, Prophet_Forecast_base, xlabel = "Date", ylabel = "Gold Close Price ($)") + ggtitle(paste0("Gold", ": Baseline Price Prediction"))

# Plot actual
p1 <- ggplot(results, aes(Date, Close_Actual, group=1)) + 
    geom_line() +
    theme_light() + ggtitle("Actual 1 Week  Price")

# Plot Predicted  
p2 <- ggplot(results, aes(Date, Close_Prediction, group=1)) + 
    geom_line() +
    theme_light() + ggtitle("Predicted 1 Week Gold Price")
# Combine Predicted and Actual Plot 
p <- p1 + 
  geom_point(mapping=p1$mapping) +
  geom_line(color='red') +
  geom_point(mapping=p2$mapping)+
  geom_line(mapping=p2$mapping, color='blue') +
  ggtitle("Predicted(blue) vs Actual(red) - 1 week forecast Baseline")
p

# Plot prophet components  
prophet_plot_components(Prophet, Prophet_Forecast_base)
```
## Buy/Sell Baseline Prophet Forecast
```{r}
results$Index <- seq(1, nrow(results), by=1)
vec <- results$Close_Prediction
x <- locate_xtrem(vec)
x <- x %>% 
  select(c('Idx', 'Status')) %>%
  rename(Index = Idx)
results <- left_join(results, x,  by="Index")
results <- results %>% 
  select(c('Date', 'Close_Prediction', 'Close_Actual', 'Status'))
results["Status"][results["Status"] == "min"] <- "Buy"
results["Status"][results["Status"] == "max"] <- "Sell"
Profit <- na.omit(results)
Profit <- Profit %>%  
  mutate(price_diff = Close_Actual - lag(Close_Actual, default = first(Close_Actual))) %>%  
  filter(Status == 'Sell')
sum(Profit$price_diff) # TWEAK OR USE ONLINE CALC
```

## GOLD FORECAST WEEKLY ROLLOVER
Idea is that if we want to forecast a week ahead, we can use last weeks external values (5 day lag) to forecast gold prices. All data will be known, but continuous forecast require weekly updates. The goal is to predict 1 week in advance, then training data gets expanded 1 week later with actual gold values, retrains model with actual data, and forecasts with 1 week lag predictors.
```{r}
# Create a new full data frame
FULL_df1 <- subset(FULL_df, Date >= as.Date("2018-01-08"))

# Lag External Regressors by 5 (1 week lag)
FULL_df1$Silver_Close.x <- lag(FULL_df1$Silver_Close.x, n=5, default = NA)
FULL_df1$DFF.x <- lag(FULL_df1$DFF.x, n=5, default = NA)
FULL_df1$WM2NS.x <- lag(FULL_df1$WM2NS.x, n=5, default = NA)
FULL_df1$hits.x <- lag(FULL_df1$hits.x, n=5, default = NA)
FULL_df1$DJIA.x <- lag(FULL_df1$DJIA.x, n=5, default = NA)
# Grab data frame to start at the beginning of the week (monday)
FULL_df1 <- subset(FULL_df1, Date >= as.Date("2018-02-12"))

# Store future dataframe (for prophet) with 1 week lag predictors 
df_future <- FULL_df1 %>%
  select(c('Date', 'Silver_Close.x', "DFF.x", "WM2NS.x" , "hits.x", "DJIA.x")) %>%
  rename(ds = Date)
```


## Gold 1 week rollover prophet forecast
```{r}
# Partition data
prophet.Train <- FULL_df1[1:1160,]
prophet.Test <- FULL_df1[1161:1164,] 
Results <- data.frame()
pred <- data.frame()
# Jumps 4 days in advance
stepsAhead <- 4
# Predicting 4 weeks of forecasts
periods_forecast <- 4
length <- seq(4, stepsAhead * periods_forecast , by=stepsAhead)

# Actual Gold Values
Gold_Results <- gold %>%
  select(c("Date","Gold_Close"))  %>%
  rename(Close_Actual = Gold_Close)
 
# Create a for loop to retrain model with a new training set
for(i in length) {
  df3 <- prophet.Train %>%
  select(c('Date', 'Gold_Close.x', 'Silver_Close.x', "DFF.x", "WM2NS.x" , "hits.x", "DJIA.x")) %>%
  rename(ds = Date, y = Gold_Close.x)
  # Add Regressors
  m_ext <- prophet(seasonality.mode = "multiplicative", daily.seasonality = FALSE, interval.width = .90)
  m_ext <- add_regressor(m_ext,  "Silver_Close.x", mode = 'multiplicative', standardize = "auto")
  m_ext <- add_regressor(m_ext,  'DFF.x', mode = 'multiplicative', standardize = "auto")
  m_ext <- add_regressor(m_ext,  'WM2NS.x', mode = 'multiplicative', standardize = "auto")
  m_ext <- add_regressor(m_ext,  'hits.x', mode = 'multiplicative', standardize = "auto")
  m_ext <- add_regressor(m_ext,  'DJIA.x', mode = 'multiplicative', standardize = "auto")
  # Fit Model
  m_ext <- fit.prophet(m_ext, df3)
  # Create future dataframe
  future <- make_future_dataframe(m_ext, periods = 6, include_history = TRUE) %>% filter(!wday(ds) %in% c(1,7))
  future <- left_join(future, df_future, by="ds")
  # Forecast with model
  forecast_weekly <- predict(m_ext, future) 
  # Return Results and organize data
  Prophet_Results <- forecast_weekly %>%
  select(c("ds","yhat")) %>%
  rename(Date = ds, Close_Prediction = yhat)
  Prophet_Results <- Prophet_Results[(nrow(prophet.Train)+1:(nrow(prophet.Train)+stepsAhead + 1)),]
  Prophet_Results <- na.omit(Prophet_Results)
  pred <- rbind(pred, Prophet_Results )
  # Re-size training and test data
  prophet.Train <- FULL_df1[1:(1160+i),]
  prophet.Test <- FULL_df1[(1161+i):(1161+i+ stepsAhead),]
  
}
# Join Predicted and Actual Results
all_results <- left_join(pred, Gold_Results,  by="Date")
all_results

# Print RMSE
RMSE <- sqrt(mean((all_results$Close_Actual - all_results$Close_Prediction)^2))
RMSE 

```
## Plot Forecast Values
```{r}
dyplot.prophet(m_ext, forecast_weekly)
prophet_plot_components(m_ext, forecast_weekly)

 p1_ext <- ggplot(all_results, aes(Date, Close_Actual, group=1)) + 
    geom_line() +
    theme_light() + ggtitle("Actual 1 Week  Price")
  
  
  p2_ext <- ggplot(all_results, aes(Date, Close_Prediction, group=1)) + 
    geom_line() +
    theme_light() + ggtitle("Predicted 1 Week Gold Price")
  
  p_ext <- p1_ext + 
    geom_point(mapping=p1_ext$mapping) +
    geom_line(color='red') +
    geom_point(mapping=p2_ext$mapping)+
    geom_line(mapping=p2_ext$mapping, color='blue') +
    ggtitle("Predicted(blue) vs Actual(red) - Rolling 1 Week - 1 Month Forecast")
  p_ext
  
  
  
```



## Buy/Sell Signal
```{r, warning=FALSE}
# Find local min and max to create buy/sell signals
all_results$Index <- seq(1, nrow(all_results), by=1)
vec <- all_results$Close_Prediction
x <- locate_xtrem(vec)
x <- x %>% 
  select(c('Idx', 'Status')) %>%
  rename(Index = Idx)
all_results <- left_join(all_results, x,  by="Index")
all_results <- all_results %>% 
  select(c('Date', 'Close_Prediction', 'Close_Actual', 'Status'))
all_results["Status"][all_results["Status"] == "min"] <- "Buy"
all_results["Status"][all_results["Status"] == "max"] <- "Sell"
Profit <- na.omit(all_results)

# Calculate profit based on buy/sell signals
Profit <- Profit %>%  
  mutate(price_diff = Close_Actual - lag(Close_Actual, default = first(Close_Actual))) %>%  
  filter(Status == 'Sell')

sum(Profit$price_diff)  # TWEAK OR USE ONLINE CALC



```


## GOLD FORECAST ROLLOVER 1 Day
Idea is that if we want to forecast 1 day ahead, we can use last weeks external values (1 day lag) to forecast gold prices. All data will be known, but continuous forecast require daily updates. The goal is to predict 1 day in advance, then training data gets expanded 1 day later with actual gold values, retrains model with actual data, and forecasts with 1 day lag predictors.
```{r}
# New dataframe with lagged predictors
FULL_df2 <- subset(FULL_df, Date >= as.Date("2018-01-08"))

# Lag External Regressors by 1 day
FULL_df2$Silver_Close.x <- lag(FULL_df2$Silver_Close.x, n=1, default = NA)
FULL_df2$DFF.x <- lag(FULL_df2$DFF.x, n=1, default = NA)
FULL_df2$WM2NS.x <- lag(FULL_df2$WM2NS.x, n=1, default = NA)
FULL_df2$hits.x <- lag(FULL_df2$hits.x, n=1, default = NA)
FULL_df2$DJIA.x <- lag(FULL_df2$DJIA.x, n=1, default = NA)
FULL_df2 <- subset(FULL_df2, Date >= as.Date("2018-02-12"))
# Store lagged variables in future dataframe for prophet
df_future <- FULL_df2 %>%
  select(c('Date', 'Silver_Close.x', "DFF.x", "WM2NS.x" , "hits.x", "DJIA.x")) %>%
  rename(ds = Date)

```



```{r}
# Partition data
prophet.Train <- FULL_df2[1:1160,]
prophet.Test <- FULL_df2[1161:1161,] 
Results <- data.frame()
pred <- data.frame()
# Jump 1 day in advance
stepsAhead <- 1
# Forecasting for 20 periods aka 20 days
periods_forecast <- 20
length <- seq(1, stepsAhead * periods_forecast , by=stepsAhead)

# Actual gold values
Gold_Results <- gold %>%
  select(c("Date","Gold_Close"))  %>%
  rename(Close_Actual = Gold_Close)
 
# Create for loop to retrain model with a new training set
for(i in length) {
  df3 <- prophet.Train %>%
  select(c('Date', 'Gold_Close.x', 'Silver_Close.x', "DFF.x", "WM2NS.x" , "hits.x", "DJIA.x")) %>%
  rename(ds = Date, y = Gold_Close.x)
  # Add regressor
  m_ext <- prophet(seasonality.mode = "multiplicative", daily.seasonality = FALSE, interval.width = .90)
  m_ext <- add_regressor(m_ext,  "Silver_Close.x", mode = 'multiplicative', standardize = "auto")
  m_ext <- add_regressor(m_ext,  'DFF.x', mode = 'multiplicative', standardize = "auto")
  m_ext <- add_regressor(m_ext,  'WM2NS.x', mode = 'multiplicative', standardize = "auto")
  m_ext <- add_regressor(m_ext,  'hits.x', mode = 'multiplicative', standardize = "auto")
  m_ext <- add_regressor(m_ext,  'DJIA.x', mode = 'multiplicative', standardize = "auto")
  #Fit Model
  m_ext <- fit.prophet(m_ext, df3)
  # Create future dataframe
  future <- make_future_dataframe(m_ext, periods = 1, include_history = TRUE) %>% filter(!wday(ds) %in% c(1,7))
  future <- left_join(future, df_future, by="ds")
  # Forecast with model
  forecast_weekly <- predict(m_ext, future) 
  
  # Return Results
  Prophet_Results <- forecast_weekly %>%
  select(c("ds","yhat")) %>%
  rename(Date = ds, Close_Prediction = yhat)
  Prophet_Results <- Prophet_Results[(nrow(prophet.Train)+1:(nrow(prophet.Train)+stepsAhead + 1)),]
  Prophet_Results <- na.omit(Prophet_Results)
  pred <- rbind(pred, Prophet_Results )
  # Resize training data
  prophet.Train <- FULL_df2[1:(1160+i),]
  prophet.Test <- FULL_df2[(1161+i):(1161+i+ stepsAhead),]
  
}

# Join Predicted and actual Results
all_results <- left_join(pred, Gold_Results,  by="Date")
all_results
# RMse
RMSE <- sqrt(mean((all_results$Close_Actual - all_results$Close_Prediction)^2))
RMSE
```
## Plot Forecast
```{r}
# FOrecast plot
dyplot.prophet(m_ext, forecast_weekly)
# Plot components
prophet_plot_components(m_ext, forecast_weekly)
# Plot actual and predicted data
 p1_ext <- ggplot(all_results, aes(Date, Close_Actual, group=1)) + 
    geom_line() +
    theme_light() + ggtitle("Actual 1 Week  Price")

 p2_ext <- ggplot(all_results, aes(Date, Close_Prediction, group=1)) + 
    geom_line() +
    theme_light() + ggtitle("Predicted 1 Week Gold Price")
 
p_ext <- p1_ext + 
    geom_point(mapping=p1_ext$mapping) +
    geom_line(color='red') +
    geom_point(mapping=p2_ext$mapping)+
    geom_line(mapping=p2_ext$mapping, color='blue') +
    ggtitle("Predicted(blue) vs Actual(red) - Rolling 1 Day - 1 Month Forecast")
 p_ext
  

```

## Buy/Sell Signal
```{r}
all_results$Index <- seq(1, nrow(all_results), by=1)
vec <- all_results$Close_Prediction
x <- locate_xtrem(vec)
x <- x %>% 
  select(c('Idx', 'Status')) %>%
  rename(Index = Idx)
all_results <- left_join(all_results, x,  by="Index")
all_results <- all_results %>% 
  select(c('Date', 'Close_Prediction', 'Close_Actual', 'Status'))
all_results["Status"][all_results["Status"] == "min"] <- "Buy"
all_results["Status"][all_results["Status"] == "max"] <- "Sell"
Profit <- na.omit(all_results)
Profit <- Profit %>%  
  mutate(price_diff = Close_Actual - lag(Close_Actual, default = first(Close_Actual))) %>%  
  filter(Status == 'Sell')
sum(Profit$price_diff) # TWEAK OR USE ONLINE CALC
```

## No Lag on Predictors (Check to see if external regressors help prediction if forecasted perfectly)
```{r}
# New dataframe with lagged predictors
FULL_df2 <- subset(FULL_df, Date >= as.Date("2018-01-08"))

# Lag External Regressors by 1 day
FULL_df2$Silver_Close.x <- FULL_df2$Silver_Close.x
FULL_df2$DFF.x <- FULL_df2$DFF.x
FULL_df2$WM2NS.x <- FULL_df2$WM2NS.x
FULL_df2$hits.x <- FULL_df2$hits.x
FULL_df2$DJIA.x <- FULL_df2$DJIA.x
FULL_df2 <- subset(FULL_df2, Date >= as.Date("2018-02-12"))
# Store lagged variables in future dataframe for prophet
df_future <- FULL_df2 %>%
  select(c('Date', 'Silver_Close.x', "DFF.x", "WM2NS.x" , "hits.x", "DJIA.x")) %>%
  rename(ds = Date)

```



```{r}
# Partition data
prophet.Train <- FULL_df2[1:1160,]
prophet.Test <- FULL_df2[1161:1161,] 
Results <- data.frame()
pred <- data.frame()
# Jump 1 day in advance
stepsAhead <- 1
# Forecasting for 20 periods aka 20 days
periods_forecast <- 20
length <- seq(1, stepsAhead * periods_forecast , by=stepsAhead)

# Actual gold values
Gold_Results <- gold %>%
  select(c("Date","Gold_Close"))  %>%
  rename(Close_Actual = Gold_Close)
 
# Create for loop to retrain model with a new training set
for(i in length) {
  df3 <- prophet.Train %>%
  select(c('Date', 'Gold_Close.x', 'Silver_Close.x', "DFF.x", "WM2NS.x" , "hits.x", "DJIA.x")) %>%
  rename(ds = Date, y = Gold_Close.x)
  # Add regressor
  m_ext <- prophet(seasonality.mode = "multiplicative", daily.seasonality = FALSE, interval.width = .90)
  m_ext <- add_regressor(m_ext,  "Silver_Close.x", mode = 'multiplicative', standardize = "auto")
  m_ext <- add_regressor(m_ext,  'DFF.x', mode = 'multiplicative', standardize = "auto")
  m_ext <- add_regressor(m_ext,  'WM2NS.x', mode = 'multiplicative', standardize = "auto")
  m_ext <- add_regressor(m_ext,  'hits.x', mode = 'multiplicative', standardize = "auto")
  m_ext <- add_regressor(m_ext,  'DJIA.x', mode = 'multiplicative', standardize = "auto")
  #Fit Model
  m_ext <- fit.prophet(m_ext, df3)
  # Create future dataframe
  future <- make_future_dataframe(m_ext, periods = 1, include_history = TRUE) %>% filter(!wday(ds) %in% c(1,7))
  future <- left_join(future, df_future, by="ds")
  # Forecast with model
  forecast_weekly <- predict(m_ext, future) 
  
  # Return Results
  Prophet_Results <- forecast_weekly %>%
  select(c("ds","yhat")) %>%
  rename(Date = ds, Close_Prediction = yhat)
  Prophet_Results <- Prophet_Results[(nrow(prophet.Train)+1:(nrow(prophet.Train)+stepsAhead + 1)),]
  Prophet_Results <- na.omit(Prophet_Results)
  pred <- rbind(pred, Prophet_Results )
  # Resize training data
  prophet.Train <- FULL_df2[1:(1160+i),]
  prophet.Test <- FULL_df2[(1161+i):(1161+i+ stepsAhead),]
  
}

# Join Predicted and actual Results
all_results <- left_join(pred, Gold_Results,  by="Date")
all_results
# RMse
RMSE <- sqrt(mean((all_results$Close_Actual - all_results$Close_Prediction)^2))
RMSE
```
## Plot Forecast
```{r}
# FOrecast plot
dyplot.prophet(m_ext, forecast_weekly)
# Plot components
prophet_plot_components(m_ext, forecast_weekly)
# Plot actual and predicted data
 p1_ext <- ggplot(all_results, aes(Date, Close_Actual, group=1)) + 
    geom_line() +
    theme_light() + ggtitle("Actual 1 Week  Price")

 p2_ext <- ggplot(all_results, aes(Date, Close_Prediction, group=1)) + 
    geom_line() +
    theme_light() + ggtitle("Predicted 1 Week Gold Price")
 
p_ext <- p1_ext + 
    geom_point(mapping=p1_ext$mapping) +
    geom_line(color='red') +
    geom_point(mapping=p2_ext$mapping)+
    geom_line(mapping=p2_ext$mapping, color='blue') +
    ggtitle("Predicted(blue) vs Actual(red) - Rolling 1 Day Forecast (No Lagged Predictors)")
 p_ext
  

```

## Buy/Sell Signal
```{r}
all_results$Index <- seq(1, nrow(all_results), by=1)
vec <- all_results$Close_Prediction
x <- locate_xtrem(vec)
x <- x %>% 
  select(c('Idx', 'Status')) %>%
  rename(Index = Idx)
all_results <- left_join(all_results, x,  by="Index")
all_results <- all_results %>% 
  select(c('Date', 'Close_Prediction', 'Close_Actual', 'Status'))
all_results["Status"][all_results["Status"] == "min"] <- "Buy"
all_results["Status"][all_results["Status"] == "max"] <- "Sell"
Profit <- na.omit(all_results)
Profit <- Profit %>%  
  mutate(price_diff = Close_Actual - lag(Close_Actual, default = first(Close_Actual))) %>%  
  filter(Status == 'Sell')
sum(Profit$price_diff) # TWEAK OR USE ONLINE CALC
```
