---
title: "Gold Prediction ADS 506 Project 2"
author: "Group 5"
date: "11/24/2022"
output: pdf_document
---

```{r setup, include=FALSE}
library(fpp2) 
library(zoo)
library(tidyr)
library(ggplot2)
library(readr)
library(dplyr)
library(rvest)
library(Quandl)
library(plotly)
library(ggplot2)
library(DT)
library(gtrendsR)
library(tidytext)
library(stringr)
library(lubridate)
#library(prophet)
library(imputeTS)
library(corrplot) 
library(caret)
#install.packages("tidytext")
#install.packages("imputeTS")
```

```{r}
#install.packages("Quandl")
#install.packages("gtrendsR")
#install.packages("tidytext")
#install.packages("imputeTS")
```

```{r}
# Load Data
#gold <- read_csv("/Users/ryan_s_dunn/Documents/USD MS-ADS/Applied Time Series Analysis ADS 506/Final Project/Gold_Nasdaq_Max.csv", 
#    col_types = cols(Date = col_date(format = "%m/%d/%Y")))
#
#unemp <- read_csv("/Users/ryan_s_dunn/Documents/USD MS-ADS/Applied Time Series Analysis ADS 506/Final #Project/UNRATE.csv", show_col_types = FALSE)
#
#M2 <- read_csv("/Users/ryan_s_dunn/Documents/USD MS-ADS/Applied Time Series Analysis ADS 506/Final #Project/WM2NS.csv", show_col_types = FALSE)
#
#DJ <- read_csv("//Users/ryan_s_dunn/Documents/USD MS-ADS/Applied Time Series Analysis ADS 506/Final Project/DJIA #(1).csv",  show_col_types = FALSE)                
#
#fed_funds <- read_csv("/Users/ryan_s_dunn/Documents/USD MS-ADS/Applied Time Series Analysis ADS 506/Final #Project/FederalFunds.csv", show_col_types = FALSE)
#
#silver <- Quandl('LBMA/SILVER')
#
#dollar_index <- read_csv("/Users/ryan_s_dunn/Documents/USD MS-ADS/Applied Time Series Analysis ADS 506/Final #Project/US Dollar Index Historical Data.csv", col_types = cols(Date = col_date(format = "%m/%d/%Y")))


gold <- read_csv("/Users/ryan_s_dunn/Documents/USD MS-ADS/Applied Time Series Analysis ADS 506/Final Project/Gold.csv", 
    col_types = cols(Date = col_date(format = "%m/%d/%Y")))

unemp <- read_csv("/Users/ryan_s_dunn/Documents/USD MS-ADS/Applied Time Series Analysis ADS 506/Final Project/UNRATE.csv", show_col_types = FALSE)

M2 <- read_csv("/Users/ryan_s_dunn/Documents/USD MS-ADS/Applied Time Series Analysis ADS 506/Final Project/WM2NS.csv", show_col_types = FALSE)

DJ <- read_csv("/Users/ryan_s_dunn/Documents/USD MS-ADS/Applied Time Series Analysis ADS 506/Final Project/DJIA.csv",  show_col_types = FALSE)                

fed_funds <- read_csv("/Users/ryan_s_dunn/Documents/USD MS-ADS/Applied Time Series Analysis ADS 506/Final Project/DFF.csv", show_col_types = FALSE)

silver <- Quandl('LBMA/SILVER')

dollar_index <- read_csv("/Users/ryan_s_dunn/Documents/USD MS-ADS/Applied Time Series Analysis ADS 506/Final Project/US Dollar Index Historical Data.csv", col_types = cols(Date = col_date(format = "%m/%d/%Y")))

Google <- read_csv("/Users/ryan_s_dunn/Documents/USD MS-ADS/Applied Time Series Analysis ADS 506/Final Project/Gold_Trend_hits.csv", show_col_types = FALSE)
```

```{r, warning=FALSE}
#edit the column names in data frames 
Price_plot <- gold %>%
  plot_ly(x = ~Date,
          type = "candlestick", 
          open = ~Open, 
          close = ~`Close/Last`, 
          high = ~High,
          low = ~Low,
          name = "price") %>%
  layout(
    xaxis = list(
      rangeselector = list(
        buttons = list(
          list(
            count = 3,
            label = "3 mo",
            step = "month",
            stepmode = "backward"),
          list(
            count = 6,
            label = "6 mo",
            step = "month",
            stepmode = "backward"),
          list(
            count = 1,
            label = "1 yr",
            step = "year",
            stepmode = "backward"),
          list(
            count = 2,
            label = "2 yr",
            step = "year",
            stepmode = "backward"),
          list(
            count = 3,
            label = "3 yr",
            step = "year",
            stepmode = "backward"),
          list(
            count = 5,
            label = "5 yr",
            step = "year",
            stepmode = "backward"),
          list(step = "all"))),
      
      rangeslider = list(visible = FALSE)),
         yaxis = list(title = "Price ($)",
                      showgrid = TRUE,
                      showticklabels = TRUE))
Volume <- select(gold, Date, Volume)
Volume$Date <- as.Date(Volume$Date , format = "%m/%d/%y")
Volume$Vol <- as.numeric(as.character(Volume$Volume)) / 1000
Volume_plot <- Volume %>%
  plot_ly(x=~Date, y=~Vol, type='bar', name = "Volume") %>%
  layout(yaxis = list(title = "Volume (Units of Thousand)"))
plot <- subplot(Price_plot, Volume_plot, heights = c(0.7,0.3), nrows=2,
             shareX = TRUE, titleY = TRUE) %>%
  layout(title = 'GC:CMX')
plot
```

#### Adjust data types where needed 
```{r}
#change data type from chr to date in the gold data set
gold$Date <- as.Date(gold$Date, format = "%m/%d/%y")
gold$DATE <- gold$Date
#change the datatype to date in the DJIA data set
DJ$DATE <- as.Date(DJ$DATE, format = "%m/%d/%y")
#indicate columns to keep in gold data set
keep_cols <-  c("Date","Close/Last", 'Volume')
gold <- gold[keep_cols] 
gold <- gold %>%
  rename(Gold_Close = 'Close/Last', Gold_Volume ='Volume')
head(gold)
```
#### Create the daily M2 rate data frame (in billions)
```{r}
#create all dates from the date column in the data frame
all_dates <- M2 %>%select(DATE) %>% 
  complete(DATE=  seq.Date(min(DATE), max(DATE), by="day"),
           ) %>% mutate(TIME=paste(year(DATE),str_pad(month(DATE), 2, pad = "0"),sep = "-"))
#join the all_dates object back onto original datafarme
M2_daily <- left_join(all_dates,M2,by="DATE")
#fill the NA values from the join with the first value from the dataset


M2_daily <- na_locf(M2_daily, option = "locf")
#M2_daily <- na.locf(M2_daily, fromLast = TRUE)


keep_cols <- c("DATE", "WM2NS")
M2_daily <- M2_daily[keep_cols]
M2_daily <- M2_daily %>%
  select(c("DATE","WM2NS")) %>%
  rename(Date = DATE)
head(M2_daily)                      
```

#### Create the daily unemployment rate data frame
```{r}
#create all dates from the date column in the dataframe
all_dates <- unemp %>%select(DATE) %>% 
  complete(DATE=  seq.Date(min(DATE), max(DATE), by="day"),
           ) %>% mutate(TIME=paste(year(DATE),str_pad(month(DATE), 2, pad = "0"),sep = "-"))
#join the all_dates object back onto original datafarme
unemp_daily <- left_join(all_dates,unemp,by="DATE")

#fill the NA values from the join with the first value from the dataset
unemp_daily <- na_locf(unemp_daily,  option = "locf")
#unemp_daily <- na.locf(unemp_daily, fromLast = TRUE)

keep_cols <- c("DATE", "UNRATE")
unemp_daily <- unemp_daily[keep_cols]
unemp_daily <- unemp_daily %>%
  select(c("DATE","UNRATE")) %>%
  rename(Date = DATE)
head(unemp_daily) 
```

#### Create the daily federal funds data frame
```{r}
#create all dates from the date column in the data frame
all_dates <- fed_funds %>%select(observation_date) %>% 
  complete(observation_date =  seq.Date(min(observation_date), max(observation_date), by="day"),
           ) %>% mutate(TIME=paste(year(observation_date),str_pad(month(observation_date), 2, pad = "0"),sep = "-"))
#join the all_dates object back onto original dataframe
ff_daily <- left_join(all_dates,fed_funds,by="observation_date")

#fill the NA values from the join with the first value from the dataset
ff_daily <- na_locf(ff_daily, option = "locf")
#ff_daily <- na.locf(ff_daily, fromLast = TRUE)

keep_cols <- c("observation_date", "DFF")
ff_daily <- ff_daily[keep_cols]
ff_daily <- ff_daily %>%
  select(c("observation_date","DFF")) %>%
  rename(Date = observation_date)
head(ff_daily) 
```
#### Create the daily Dow Jones IA data frame
```{r}

#create all dates from the date column in the data frame
all_dates <- DJ %>%select(DATE) %>% 
  complete(DATE =  seq.Date(min(DATE), max(DATE), by="day"),
           ) %>% mutate(TIME=paste(year(DATE),str_pad(month(DATE), 2, pad = "0"),sep = "-"))
#join the all_dates object back onto original datafarme
dj_daily <- left_join(all_dates, DJ ,by="DATE")

#fill the NA values from the join with the first value from the dataset
dj_daily <- na_locf(dj_daily,  option = "locf")
#dj_daily <- na.locf(dj_daily, fromLast = TRUE)

keep_cols <- c("DATE", "DJIA")
dj_daily <- dj_daily[keep_cols]
dj_daily <- dj_daily %>%
  select(c("DATE", "DJIA")) %>%
  rename(Date = DATE)
head(dj_daily) 
```

#### Silver

```{r}
# add in Silver
silver <- silver[,c('Date', 'USD')]
silver <- silver %>%
  rename(Silver_Close = 'USD')
Metals <- inner_join(silver, gold, by="Date")
Metals %>% arrange(mdy(Metals$Date))
head(Metals)
```

### Dollar Index
```{r}
# Add in Dollar Index
dollar_index <- dollar_index %>%
  select(c("Date","Price")) %>%
  rename(DXY = Price)
#dollar_index$Date <- as.Date(dollar_index$Date, format = "%m/%d/%y")
```

#### Google trends
```{r}
#trends <- gtrends(keyword = "Gold", onlyInterest = TRUE, time = "today+5-y")


#create all dates from the date column in the data frame
all_dates <- Google %>%select(Date) %>% 
  complete(Date =  seq.Date(min(Date), max(Date), by="day"),
           ) %>% mutate(TIME=paste(year(Date),str_pad(month(Date), 2, pad = "0"),sep = "-"))
#join the all_dates object back onto original datafarme
Google_daily <- left_join(all_dates, Google ,by="Date")

#fill the NA values from the join with the first value from the dataset
Google_daily <- na_locf(Google_daily,  option = "locf")


keep_cols <- c("Date", "Gold: (Worldwide)")
Google_daily <- Google_daily[keep_cols]
Google_daily <- Google_daily %>%
  select(c("Date", "Gold: (Worldwide)")) %>%
  rename(hits = 'Gold: (Worldwide)')


Google_daily %>%  
  plot_ly(type='scatter',x=~Date, y=~hits, mode = 'lines', name = "Google Search Trends") %>%
  layout(title = paste0("Interest over Time: ", "Gold"), yaxis = list(title = "hits"))

trends_daily <- Google_daily
# Hits verse Gold 
trends_daily %>%
  left_join(gold, by = "Date") %>%
  select(one_of(c("Date", "hits", "Gold_Close"))) %>%
  drop_na() %>%
  ggplot(aes(hits, Gold_Close)) + geom_point(color="red") + geom_smooth(method = 'loess') +
  labs(title =paste0("Gold",": Relationship between World Interest (Hits) and Close Price (Gold)"))

```


#### Merge data frames and create the final dataframe for EDA
```{r}
#merge the dataframes together on dates
full_df <- left_join(Metals, dollar_index, by="Date")
full_df <- left_join(full_df, unemp_daily, by="Date")
full_df <- left_join(full_df, ff_daily, by="Date")
full_df <- left_join(full_df, M2_daily, by="Date")
full_df <- left_join(full_df, Google_daily, by="Date")
full_df <- left_join(full_df, dj_daily, by="Date")
full_df <- left_join(full_df, trends_daily, by="Date")
full_df$Gold_Volume <- na_ma(full_df$Gold_Volume, k=1, weighting = "simple")

keep_cols <- c("Date", "Silver_Close", "Gold_Close","Gold_Volume", "DXY", "UNRATE", "DFF", "WM2NS" ,"hits.x",     "DJIA")     

full_df <- full_df[keep_cols]
full_df <- full_df %>%
  rename(hits = hits.x)

full_df <- full_df %>% arrange(mdy(full_df$Date))
head(full_df)
```

```{r}
# Check for any NA Values
cbind(
   lapply(
     lapply(full_df, is.na)
     , sum)
   )
```
#### Subset the data frame with values from 2018 forward
```{r}
full_df <- full_df[full_df$Date >= "2018-01-01",]

#Replace NA with the mose recent value
full_df$WM2NS <- zoo::na.fill(full_df$WM2NS, "extend")
full_df$hits <- zoo::na.fill(full_df$hits, "extend")
full_df$UNRATE <- zoo::na.fill(full_df$UNRATE, "extend")

#removes the data that is being pulled in daily for silver - since the .csv's are stagnet
#full_df <- na.omit(full_df)
full_df$Date <- as.Date(full_df$Date)
full_df <- full_df %>% arrange(ymd(full_df$Date))
head(full_df)
```

```{r}
# Check for any NA Values
cbind(
   lapply(
     lapply(full_df, is.na)
     , sum)
   )
```

#### Create the correlation matrix for EDA 
```{r}
#correlation matrix of all predictors to gold 
#corr_fields <- c("Gold_Close", 'Gold_Volume', 'Silver_Close', 'hits', "UNRATE","DFF","WM2NS", "DJIA")
#corr_fields <- c("Gold_Close", 'Gold_Volume', 'Silver_Close',  "UNRATE","DFF","WM2NS","DJIA")
corr_fields <- c("Gold_Close", "Silver_Close","Gold_Volume", "DXY", "UNRATE", "DFF", "WM2NS" ,"hits", "DJIA")     
full_df$Gold_Volume <- as.numeric(full_df$Gold_Volume)
full_df$WM2NS <- as.numeric(full_df$WM2NS)
full_df$hits <- as.numeric(full_df$hits)

full_df$Gold_Volume <- na_ma(full_df$Gold_Volume, k=1, weighting = "simple")
full_df$WM2NS <- na_ma(full_df$WM2NS, k=1, weighting = "simple")
full_df$hits <- na_ma(full_df$hits, k=1, weighting = "simple")
```
```{r}
corr_df <- full_df[corr_fields]
corr_matrix = cor(corr_df)
#display first row of correlation matrix
corr_matrix[,1]
```

```{r}
#develop the corrplot correlation matrix
corrplot(corr_matrix)
```
#### Normalize variables for EDA
```{r}
#normalize variables to view scaled relationships and add into full_df
full_df_Norm <- full_df
full_df_Norm[-1] <- lapply(full_df_Norm[-1], scale)
```

#### EDA with non-normalized varialbes
````{r}
#scatterplot of gold & each external variable

ggplot(data = full_df) +
  geom_point(mapping = aes(x = Gold_Close,
                            y = UNRATE, color = Date)) + 
  ggtitle("Scatterplot of Gold Close Price and Unemployment Rate (Date Color)")

ggplot(data = full_df) +
  geom_point(mapping = aes(x = Gold_Close,
                            y = DFF, color = Date)) + 
  ggtitle("Scatterplot of Gold Close Price and Federal Funds Rate (Date Color)")

ggplot(data = full_df) +
  geom_point(mapping = aes(x = Gold_Close,
                            y = WM2NS, color = Date)) + 
  ggtitle("Scatterplot of Gold Close Price and M2 Money Supply (Date Color)")

ggplot(data = full_df) +
  geom_point(mapping = aes(x = Gold_Close,
                            y = DJIA, color = Date)) + 
  ggtitle("Scatterplot of Gold Close Price and DJIA (Date Color)")

ggplot(data = full_df) +
  geom_point(mapping = aes(x = Gold_Close,
                            y = Gold_Volume, color = Date)) + 
  ggtitle("Scatterplot of Gold Close Price and Gold Volume (Date Color)")

ggplot(data = full_df) +
  geom_point(mapping = aes(x = Gold_Close,
                            y = Silver_Close, color = Date)) + 
  ggtitle("Scatterplot of Gold_Close and Silver_Close (Date Color)")

ggplot(data = full_df) +
  geom_point(mapping = aes(x = Gold_Close,
                            y = hits, color = Date)) + 
  ggtitle("Scatterplot of Gold_Close and Hits (Date Color)")
```
#### Create dataframe for Normalize EDA (on same axis)
```{r}
#create time series objects for each of the normalized data elements 
gold_ts_norm <- ts(full_df_Norm$Gold_Close, start = c(2018, 01), frequency = 252)
unemp_ts_norm <- ts(full_df_Norm$UNRATE, start = c(2018, 01), frequency = 252)
ff_ts_norm <- ts(full_df_Norm$DFF, start = c(2018, 01), frequency = 252)
M2_ts_norm <- ts(full_df_Norm$WM2NS, start = c(2018, 01), frequency = 252)
dj_ts_norm <- ts(full_df_Norm$DJIA, start = c(2018, 01), frequency = 252)
silver_ts_norm <- ts(full_df_Norm$Silver_Close, start = c(2018, 01), frequency = 252)
hits_ts_norm <- ts(full_df_Norm$hits, start = c(2018, 01), frequency = 252)
gold_vol_ts_norm <- ts(full_df_Norm$Gold_Volume, start = c(2018, 01), frequency = 252)
```

```{r}
#line chart of all indicators
autoplot(gold_ts_norm, main = "5-Year Trend of Gold and Unemployment Rate", xlab = "Date") + 
  autolayer(unemp_ts_norm) + theme_classic()

autoplot(gold_ts_norm, main = "5-Year Trend of Gold and Federal Funds Rate", xlab = "Date") + 
  autolayer(ff_ts_norm) + theme_classic()

autoplot(gold_ts_norm, main = "5-Year Trend of Gold Close and M2", xlab = "Date") + 
  autolayer(M2_ts_norm) + theme_classic()

autoplot(gold_ts_norm, main = "5-Year Trend of Gold and DJIA", xlab = "Date") + 
  autolayer(dj_ts_norm) + theme_classic()

autoplot(gold_ts_norm, main = "5-Year Trend of Gold and Silver Close", xlab = "Date") + 
  autolayer(silver_ts_norm) + theme_classic()

autoplot(gold_ts_norm, main = "5-Year Trend of Gold and Hits", xlab = "Date") + 
  autolayer(hits_ts_norm) + theme_classic()

autoplot(gold_ts_norm, main = "5-Year Trend of Gold and Gold Volume", xlab = "Date") + 
  autolayer(gold_vol_ts_norm) + theme_classic()

autoplot(ts(full_df$Gold_Close, frequency = 252))

#full normalized variable plot 
autoplot(gold_ts_norm) +
  autolayer(unemp_ts_norm) +
  autolayer(ff_ts_norm) + 
  autolayer(M2_ts_norm) +
  autolayer(dj_ts_norm) +
  autolayer(silver_ts_norm) +
  autolayer(hits_ts_norm) + theme_classic()
  
  
  
```

#### Boxplot to identify outlier variables 
```{r}
#boxplot of all varialbes  
ggplot(corr_df, aes(x=Gold_Close)) +
  geom_boxplot() + ggtitle("Gold Boxplot") + xlab("Gold Closing Price")

ggplot(corr_df, aes(x=UNRATE)) +
  geom_boxplot() + ggtitle("Unemployement Rate Boxplot") + xlab("Unemployement Rate")

ggplot(corr_df, aes(x=DFF)) +
  geom_boxplot() + ggtitle("Federal Funds Rate Boxplot") + xlab("Federal Funds Rate")

ggplot(corr_df, aes(x=WM2NS)) +
  geom_boxplot() + ggtitle("M2 Money Supply Boxplot") + xlab("M2 Total Money Supply")

ggplot(corr_df, aes(x=DJIA)) +
  geom_boxplot() + ggtitle("DJIA Boxplot") + xlab("Dow Jones Industrial Average")

ggplot(corr_df, aes(x=Gold_Volume)) +
  geom_boxplot() + ggtitle("Gold Volume Boxplot") + xlab("Gold Volume")

ggplot(corr_df, aes(x=Silver_Close)) +
  geom_boxplot() + ggtitle("Silver Close Boxplot") + xlab("Silver Close")

ggplot(corr_df, aes(x=hits)) +
  geom_boxplot() + ggtitle("Hits Boxplot") + xlab("Hits")
```

### Full entire Dataset

```{r}
FULL_df  <- left_join(full_df, full_df_Norm, by="Date")
FULL_df %>% arrange(mdy(FULL_df$Date))

FULL_df$difference <- c(NA, diff(FULL_df$Gold_Close.x))
FULL_df$diff_boolean <- ifelse(FULL_df$difference >= 0,1,0)
head(FULL_df)
```
### Partition the data into training and testing
```{r}
#create the time series object for the FULL_df
ts_FULL_df <- ts(data=FULL_df$Gold_Close.x, start = c(2018,01), frequency = 252)

train_df <- window(ts_FULL_df, start = c(2018,01), end = c(2022,186)) #186 is Oct 31
test_df <-  window(ts_FULL_df, start = c(2022,187))
  
#validate the train and test layers connect
autoplot(train_df) + autolayer(test_df) + 
  labs(title = "2022 - 2023 Gold Close Price (Train & Test)",
       x = "Time", y = "Gold Close Price") 

head(train_df) 
```

#### Auto ARIMA Model 1 (Gold Close Price Only ) - Proof of Random Walk
```{r}
#ARIMA Model #1
auto_model <- auto.arima(train_df)
summary(auto_model)

#investigate the potential parameters of the ARIMA model
acf(train_df, lag.max = 252)
pacf(train_df, lag.max = 252)
```
```{r}
#fit to the test data
sqrt(mean(auto_model$residuals^2))

accuracy(forecast(auto_model, h = 15), test_df)

gold_pred <- forecast(auto_model, h = 15)
#gold_pred

autoplot(train_df) + autolayer(test_df) + autolayer(gold_pred, alpha = 0.5) +coord_cartesian(c(2022,2023))
```
 

#### Logistic Regression Model #1 (All Predictors)
```{r}
#logistic regression model # 1

#columns to keep
keep_cols <- c("Date","Gold_Close.x","Silver_Close.x","DXY.x","UNRATE.x","DFF.x", "hits.x","DJIA.x","WM2NS.x", "diff_boolean")
lr_df <- FULL_df[keep_cols]
head(lr_df)
```
```{r}
#create train and test
lr_train <- lr_df[lr_df$Date <= "2022-10-31",]
lr_actual <- lr_df[lr_df$Date > "2022-10-31",]
lr_actual_gold <- lr_df[lr_df$Date > "2022-10-31",]
```


```{r}
#create the logistic regression model with all predictors 
set.seed(100)
gold_lr_model <- glm(diff_boolean ~  Silver_Close.x + DXY.x+ UNRATE.x + DFF.x + hits.x + DJIA.x + WM2NS.x , lr_train,
                    family = "binomial")

#summarize the model output
summary(gold_lr_model)
gold_lr_model
```

```{r}
#generate the predictions of the lr model
gold_pred_prob <- predict(gold_lr_model, newdata = lr_actual, type = "response")
gold_pred_df <- cbind(lr_actual, gold_pred_prob)

#identify columns to keep and create the gold boolean flag
eval_cols <- c("Date","diff_boolean","gold_pred_prob")
eval_df <- gold_pred_df[eval_cols]
eval_df$gold_boolean <- ifelse(eval_df$gold_pred_prob >= .50,1,0)

eval_df <- cbind(eval_df, lr_actual_gold$Gold_Close.x)
eval_df
```

```{r}
#create the confusion matrix for the LR model with all predictors
confusionMatrix(data = as.factor(eval_df$diff_boolean), reference = as.factor(eval_df$gold_boolean), positive = "1")
```

#### Logistic Regresson Model #2 (with highly correlated features)
```{r}
#create the data frame of values with high correlation 
high_corr_feat <- c("Date","Silver_Close.x","WM2NS.x","hits.x","DFF.x","diff_boolean")
lr_df2 <- FULL_df[high_corr_feat]
#head(lr_df2)
```

```{r}
#create train and test sets
lr_train2 <- lr_df2[lr_df2$Date <= "2022-10-31",]
lr_actual2 <- lr_df2[lr_df2$Date > "2022-10-31",]
```

```{r}
#create the logistic regression model with high only correlated predictors
set.seed(100)
#gold_lr_model2 <- glm(diff_boolean ~  Silver_Close.x  + DFF.x + hits.x + DJIA.x + WM2NS.x , lr_train2,
#                    family = "binomial")
gold_lr_model2 <- glm(diff_boolean ~  Silver_Close.x   + WM2NS.x , lr_train2,
                    family = "binomial")
#summarize the model output
summary(gold_lr_model2)
gold_lr_model2
```

```{r}
#generate the predictions of the lr_model2
gold_pred_prob2 <- predict(gold_lr_model2, newdata = lr_actual2, type = "response")
gold_pred_df2 <- cbind(lr_actual2, gold_pred_prob2)

#identify columns to keep and create the gold boolean flag
eval_cols2 <- c("Date","diff_boolean","gold_pred_prob2")
eval_df2 <- gold_pred_df2[eval_cols2]
eval_df2$gold_boolean <- ifelse(eval_df2$gold_pred_prob >= .50,1,0)

eval_df2
```
The Logistic Regression model with only highly correlated features predicts all positive cases for the gold_boolean flag
```{r}
#return the distinct values of gold Boolean
n_distinct(eval_df2$gold_boolean)
```









